Linear Variation and an Optimization of Algorithms for Connected Components Labeling in a Binary Image

### 2016-04-20 paper/paper.tex
Some proofreading, make conclusion more concrete.

@@

\begin{abstract}
The linear variation is a topological characteristic of the function of two
variables. The problem of linear variation computing can be reduced to the
problem of counting connected components in a binary image with eight-connected
connectivity.
%The new hybrid method for the problem is presented.
%The method is essentially a modification of a two-scan algorithm for labeling
%that groups the pixels into $2 \times 2$ cells.
%This is usually done through connected components labeling, and many approaches
%for that are known. This paper presents an efficient way to convert the initial
%binary image with 8-connectivity into a condensed %(non-binary) image with new
%connectivity that is 2 times smaller in each dimension than the initial image.
%Most approaches for connected components labeling are still valid for the new
%image. As time and memory consumptions of conventional approaches usually
%depend linearly on the number of pixels, running them on the condensed image
%can be up to 4 times more efficient than running on the initial image.
The method is essentially a modification of some known raster algorithms for
labeling that groups pixels into $2 \times 2$ cells. A performance benchmark of
the proposed  method applied to some approaches on noise images is provided.

@@

Continuous functions of two variables have a set of topological characteristics
referred to as linear variations. Kronrod\cite{mak2} introduced the notion of a
regular component of the level set of a continuous function of two variables.
The simplest topological characteristic in linear variation theory is the number
of regular components for all level sets of a function. In \cite{mak3,mak4,mak5}
was discussed the necessity of the using a linear variation in the image
restoration theory.

The problem of computing the linear variation in discrete case is equivalent 
to the problem of counting the number of connected components (CC) in a binary
image with 8-connectivity. This is usually done through CC labeling, and many
approaches for that are known\cite{hechao}. We came to the labeling problem
while searching the algorithm for linear variation computing. Of course, the
labeling problem also is well known in the digital image processing theory.

@@

It is worth noting that our method exploits an important property of
8-connectivity that does not hold for 4-connectivity, so it is not applicable
for the latter case.

@@

\texten{
We will use the fact that for the case of 8-connectivity all $4$ pixels of any
$2 \times 2$ square are adjacent to each other, and so are contained in the same
CC and should have same labels upon completion of the labeling algorithm. So
instead of labeling single pixels we can label $2 \times 2$ groups of pixels, or
``cells''.
}\textru{
Наблюдение: с 8-связностью в любом квадратике $2 \times 2$ любые два пикселя
находятся рядом. Значит если это пара объектных пикселей, то они находятся в
одной КС, и значит каждый такой квадратик пересекается не более чем с одной КС.
Поэтому вместо маркировки отдельных пикселей можно маркировать такие квадраты.
Назовём их клетками.
}

\texten{
Consider a $2N \times 2M$ binary image\footnote{
  For convenience in this paper we consider only input images with even size
  in both dimensions. If this is not the case, one can easily append a row or
  a column of background pixels to the image.
}
specified by a binary predicate $b(x, y)$, returning $1$ in case the pixel at
the given coordinates is black (object pixel), and $0$ otherwise (background
pixel). For convenience we use 0-indexation throughout the paper, so the topmost
row has row index $y=0$, and the leftmost column has column index $x=0$.
}\textru{
Рассмотрим бинарное изображение размера $2N \times 2M$\footnote{
  Для удобства мы рассматриваем только изображения с чётной высотой и шириной.
  В общем случае при необходимости можно добавить строку или столбец с фоновыми
  пикселями.
},
заданное предикатом $b(x, y)$, который возвращает $1$, если переданы координаты
объектного пикселя, и $0$ иначе.
Мы везде будем использовать $0$-индексацию: верхняя строка и левый столбец имеют
номер $0$.
}

@@

\texten{
We need also to define connectivity function for the condensed image.
We cannot actually just use 8-connectivity, as whether two cells are adjacent
cells of one CC depends not only on their coordinates, but also on their
configuration. See \autoref{fig:connectivity:examples} for examples.
}\textru{
Ещё нужно дать определение связности для сжатого изображения. Мы уже не можем
использовать 8-связность: теперь нельзя сказать, связаны ли две клетки, не
проанализировав их цвета. Смотри \autoref{fig:connectivity:examples}.
}

@@

\texten{
However, for every variant of relative placement of two nearby cells there is a
mask of pixels of each cells that are important for direct connectivity of this
pair of cells. So two cells are considered directly connected, if they are
nearby in terms of 8-connectivity and each of them contains at least one object
pixel in the mask of important pixels corresponding to relative placement of the
cells. The masks are shown in \autoref{fig:connectivity:masks}.
}\textru{
Тем не менее, можно заметить, что анализировать можно не все биты цвета, а
только некоторую маску тех бит, которые важны для связности двух конкретных
клеток. Эти маски показаны на \autoref{fig:connectivity:masks}.
}

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 2.5) {};
      \node[cell] at (2.5, 0.5) {};
      \node[graybox] at (1, 2) {3};
      \node[graybox] at (2, 1) {0};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 2.5) {};
      \node[cell] at (0.5, 0.5) {};
      \node[graybox] at (0, 2) {2};
      \node[graybox] at (1, 2) {3};
      \node[graybox] at (0, 1) {0};
      \node[graybox] at (1, 1) {1};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 0.5) {};
      \node[cell] at (2.5, 2.5) {};
      \node[graybox] at (1, 1) {1};
      \node[graybox] at (2, 2) {2};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.25\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 0.5) {};
      \node[cell] at (2.5, 0.5) {};
      \node[graybox] at (1, 0) {3};
      \node[graybox] at (1, 1) {1};
      \node[graybox] at (2, 0) {2};
      \node[graybox] at (2, 1) {0};
    \end{tikzpicture}
  \end{subfigure}
  \caption{Masks of important pixels for all four cases of relative placement
    of two nearby cells. Those pixels are filled in gray, with their numbers
    within cell specified.}
  \label{fig:connectivity:masks}
\end{figure}

\texten{
The idea is that if we naturally encode mask pixels as 4-bit numbers in the
way similar to how we defined $c$, we can express the predicate denoting whether
two nearby cells
are directly connected efficiently with just two bitwise and one logical AND
operations.
As an example, \autoref{fig:connectivity:code} compares possible codes in \C
checking if the labels of the current pixel or cell and the one right on top of
it should be same for classic 8-connectivity and for the new connectivity. Note
that as bitwise operations are usually relatively cheap, we added only a little
complication compared to decreasing the number of pixels to process by the
factor of 4.
}\textru{
Идея в том, что с нашим естественным способом кодирования конфигурации клеток в
4-разрядных числах, предикат, определяющий связность двух соседных клеток,
может быть представлен всего двумя битовыми и одной логической операцией И.
В качестве примера, на \autoref{fig:connectivity:code} приведено сравнение
возможных фрагментов программы на языке \C\hspace{-0.8em}, проверяющих, связан
ли текущий пиксель (или клетка) с пикселем (или клеткой) над ним (ней).
Стоит отметить, что хотя конструкция усложнилась, битовые операции как правило
очень эффективны, и это усложнение, скорее всего, незначительно по сравнению с
уменьшением количества рассматриваемых позиций в 4 раза.
}

@@

@@

\texten{
Also to avoid repeating ``pixel or cell'' we will just use the word ``pixel''
meaning the appropriate raster for non-condensed and condensed images. When we
mention neighbors of some pixel, we mean pixels that are directly connected to
it in terms of appropriate connectivity.
}\textru{
По аналогии, чтобы избежать повторения ``пиксель или клетка'' мы будет просто
использовать слово ``пиксель'', имея в виду растр входного изображения.
Под соседями пикселя будем понимать пиксели, прямо связанные с ним в терминах
соответствующей связности.
}

@@

\texten{
Also we have to mention that for performance reasons the Union-Find internal
structures are allocated ahead to have the size of at least $HW$ to guarantee
that each object can be easily associated with unique starting set given only
its coordinates. So memory consumption of SUF is linear in the number of pixels.
}\textru{
Также необходимо отметить, что для лучшей производительности служебные массивы
структуры Union-Find заранее выделяются размера как минимум $HW$, чтобы
обеспечить простоту соответствия между пикселями изображения и элементами в
системе множеств.
Таким образом, SUF использует линейное от числа пикселей количество памяти
$O(HW)$.
}

@@

\texten{
Note that, while SUF2 is intended for CC counting, its memory
consumption is only $O(W)$, which is significant for embedded systems and
other resource-tight applications.
}\textru{
Отметим, что SUF2 решает задачу подсчёта СК, используя всего $O(W)$
дополнительной памяти.
}

@@

\texten{
This modification of SUF was implemented to compare the optimization we present
with another optimization suggested by Wu, Otoo and Suzuki\cite{wuotoo}, namely
decision tree. Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not
applicable for condensed image, as it exploits similar observations on
8-connectivity.
}\textru{
Эта модификация SUF была добавлена чтобы сравнить нашу оптимизацию с
одной из оптимизаций, предложенных Wu, Otoo и Suzuki\cite{wuotoo}, а именно с
деревом принятия решения.
К сожалению, DTSUF (Decision Tree Scan with Union-Find) не применим для сжатого
изображения, так как использует свойства 8-связности, которыми наша новая
связность не обладает.
}

@@

\texten{
For memory-tight applications such as embedded systems it can be useful to
utilise the suggested condensation in a lazy way, so that without actually
constructing the condensed image we use a function (``view'') calculating color
of each cell with 4 lookups to the original binary image. This is probably most
interesting when combined with SUF2 algorithm, as it is a way to win some time
while avoiding memory footprint increase. However this lazy approach increases
the number of image lookups.
% So, according to our benchmark, this combination is efficient only for dense
% ($>30\%$) noise.
}\textru{
В некоторых случаях может быть более приемлемо использовать сжатие неявно, то
есть вместо построения 
}

@@

So benchmarks were run for several values of $d$ to achieve a picture of how the
speedup depends on the number of object pixels in the image.

It is worth noting again that for ``2x2 copy'' optimization, the apparent time
overhead of constructing the extra condensed image  was included into the total
run time, since we measured the total time between calling the CC counting
subroutine, supplied with only the original image, and returning back from
it.

@@

The overall benchmark line plot of all algorithms and optimizations is
presented on \Cref{fig:overall}.

@@

The worst-case and best-case speedups are shown on the \autoref{tab:speedup}.
Here we see that while ``2x2 view'' optimization is more memory-efficient than
``2x2 copy'', it can slow down the method in some situations, and generally it
does not give as much performance advantage.

@@

\begin{table}
  \centering
  \begin{tabular}{l|c|c|c|c}
    Algorithm & worst speedup & worst speedup density & best speedup & best
    speedup density \\
    \hline
    \input{tables/speedup}
  \end{tabular}
  \caption{The worst and the worst speedups of the considered optimizations}
  \label{tab:speedup}
\end{table}

@@

The problem of linear variation computing can be reduced to the problem of
counting connected components in a binary image with eight-connected
connectivity. This paper presents an efficient way to convert the initial binary
image with 8-connectivity into a condensed (non-binary) image with new
connectivity that is 2 times smaller in each dimension than the initial image.
Most approaches for connected components labeling are still valid for the new
image. As time and memory consumptions of conventional approaches usually depend
linearly on the number of pixels, running them on the condensed image can be up
to 4 times more efficient than running on the initial image. The method is
essentially a modification of some known raster algorithms for labeling that
groups pixels into $2 \times 2$ cells. A performance benchmark of the proposed
method applied to some approaches on noise images is provided. It is shown at
least up to 2 times more quick results could be received.

>>>>

\usepackage{braket}

@@

\begin{abstract}
Linear variation is a topological characteristic of a function of two
variables. The problem of linear variation computing can be reduced to the
problem of counting connected components in a binary image with eight-connected
connectivity.
The proposed method is essentially a modification of some known raster
algorithms for connected components labeling that groups pixels into $2 \times
2$ cells. A performance benchmark of the proposed method applied to some
approaches on random images is provided.

@@

Continuous functions of two variables have a set of topological characteristics
referred to as linear variations. Kronrod\cite{mak2} introduced the notion of
regular component of the level set of continuous function of two variables.
The simplest topological characteristic in linear variation theory is the number
of regular components for all level sets of a function. Works
\cite{mak3,mak4,mak5} address the necessity of linear variation in image
restoration theory.

The problem of computing linear variation in discrete case is equivalent to the
problem of counting the number of connected components (CC) in a binary image
with 8-connectivity. This is usually done through CC labeling, and many
approaches for that are known\cite{hechao}. We came to the labeling problem
while searching an efficient algorithm algorithm for linear variation
computation. Of course, the labeling problem also is well known in the digital
image processing theory.

@@

It is worth noting that our method exploits an important property of
$8$-connectivity that does not hold for $4$-connectivity, so it is not
applicable in the latter case.

@@

\texten{
We will use the fact that in case of 8-connectivity all $4$ pixels of any $2
\times 2$ square are adjacent to each other, and so are contained in same CC and
should have same labels upon completion of the labeling algorithm. So instead of
labeling single pixels we can label $2 \times 2$ groups of pixels, or ``cells''.
}\textru{
Наблюдение: с 8-связностью в любом квадратике $2 \times 2$ любые два пикселя
находятся рядом. Значит если это пара объектных пикселей, то они находятся в
одной КС, и значит каждый такой квадратик пересекается не более чем с одной КС.
Поэтому вместо маркировки отдельных пикселей можно маркировать такие квадраты.
Назовём их клетками.
}

\texten{
Consider a $2N \times 2M$ binary image\footnote{
  For convenience in this paper we consider only input images with even size
  in both dimensions. If this is not the case, one can easily append a row or
  a column of background pixels to the image.
}
specified by binary predicate $b(x, y)$, returning $1$ in case the pixel at the
given coordinates is an object pixel, and $0$ otherwise (background pixel). For
convenience we use 0-indexation throughout the paper, so the topmost row is
associated with row index $y=0$, and the leftmost column is associated with
column index $x=0$.
}\textru{
Рассмотрим бинарное изображение размера $2N \times 2M$\footnote{
  Для удобства мы рассматриваем только изображения с чётной высотой и шириной.
  В общем случае при необходимости можно добавить строку или столбец с фоновыми
  пикселями.
},
заданное предикатом $b(x, y)$, который возвращает $1$, если переданы координаты
объектного пикселя, и $0$ иначе.
Мы везде будем использовать $0$-индексацию: верхняя строка и левый столбец имеют
номер $0$.
}

@@

\texten{
We need also to define a connectivity function for the condensed image.
We cannot actually just use 8-connectivity, as whether two cells are adjacent
cells of one CC depends not only on their coordinates, but also on their
configuration. See \autoref{fig:connectivity:examples} for examples.
}\textru{
Ещё нужно дать определение связности для сжатого изображения. Мы уже не можем
использовать 8-связность: теперь нельзя сказать, связаны ли две клетки, не
проанализировав их цвета. Смотри \autoref{fig:connectivity:examples}.
}

@@

\texten{
However, for every possible case of relative placement of two nearby cells there
is a mask of pixels of each cells that are important for direct connectivity of
this pair of cells. So two cells are considered directly connected, if they are
nearby in terms of 8-connectivity and each of them contains at least one object
pixel in the mask of important pixels corresponding to relative placement of the
cells. The masks are shown in \autoref{fig:connectivity:masks}.
}\textru{
Тем не менее, можно заметить, что анализировать можно не все биты цвета, а
только некоторую маску тех бит, которые важны для связности двух конкретных
клеток. Эти маски показаны на \autoref{fig:connectivity:masks}.
}

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 2.5) {};
      \node[cell] at (2.5, 0.5) {};
      \node[graybox] at (1, 2) {3};
      \node[graybox] at (2, 1) {0};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 2.5) {};
      \node[cell] at (0.5, 0.5) {};
      \node[graybox] at (0, 2) {2};
      \node[graybox] at (1, 2) {3};
      \node[graybox] at (0, 1) {0};
      \node[graybox] at (1, 1) {1};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 0.5) {};
      \node[cell] at (2.5, 2.5) {};
      \node[graybox] at (1, 1) {1};
      \node[graybox] at (2, 2) {2};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 0.5) {};
      \node[cell] at (2.5, 0.5) {};
      \node[graybox] at (1, 0) {3};
      \node[graybox] at (1, 1) {1};
      \node[graybox] at (2, 0) {2};
      \node[graybox] at (2, 1) {0};
    \end{tikzpicture}
  \end{subfigure}
  \caption{Masks of important pixels for all four cases of relative placement
    of two nearby cells. Those pixels are filled in gray, with their numbers
    within cell specified.}
  \label{fig:connectivity:masks}
\end{figure}

@@

\texten{
The idea is that if we naturally encode mask pixels as 4-bit numbers in the
way similar to how we defined $c$, we can express the predicate denoting whether
two nearby cells
are directly connected efficiently with just two bitwise and one logical AND
operations.
As an example, \autoref{fig:connectivity:code} compares possible codes in \C
checking if the labels of the current pixel or cell and the one right on top of
it should be same for classic 8-connectivity and for the new connectivity. Note
that as bitwise operations are usually relatively cheap, we added only a little
complication compared to the decrease of the number of pixels to process by the
factor of 4.
}\textru{
Идея в том, что с нашим естественным способом кодирования конфигурации клеток в
4-разрядных числах, предикат, определяющий связность двух соседных клеток,
может быть представлен всего двумя битовыми и одной логической операцией И.
В качестве примера, на \autoref{fig:connectivity:code} приведено сравнение
возможных фрагментов программы на языке \C\hspace{-0.8em}, проверяющих, связан
ли текущий пиксель (или клетка) с пикселем (или клеткой) над ним (ней).
Стоит отметить, что хотя конструкция усложнилась, битовые операции как правило
очень эффективны, и это усложнение, скорее всего, незначительно по сравнению с
уменьшением количества рассматриваемых позиций в 4 раза.
}

@@

\texten{
Also to avoid repeating ``pixel or cell'' we will just use the word ``pixel''
meaning the appropriate raster for non-condensed and condensed images. When
mentioning neighbors of some pixel, we mean pixels that are directly connected
to it in terms of appropriate connectivity.
}\textru{
По аналогии, чтобы избежать повторения ``пиксель или клетка'' мы будет просто
использовать слово ``пиксель'', имея в виду растр входного изображения.
Под соседями пикселя будем понимать пиксели, прямо связанные с ним в терминах
соответствующей связности.
}

@@

\texten{
Also we have to mention that for performance reasons the Union-Find internal
structures are allocated ahead to have the size of at least $HW$ to guarantee
that each object can be easily associated with unique starting set given only
its coordinates. So memory consumption of SUF is linear in the number of pixels
($O(HW)$), while its amortized time complexity is ``almost'' linear
($O(HW\alpha(HW))$).
}\textru{
Также необходимо отметить, что для лучшей производительности служебные массивы
структуры Union-Find заранее выделяются размера как минимум $HW$, чтобы
обеспечить простоту соответствия между пикселями изображения и элементами в
системе множеств.
Таким образом, SUF использует линейное от числа пикселей количество памяти
$O(HW)$.
}

@@

\texten{
Note that, while SUF2 is intended for CC counting, its memory
consumption is only $O(W)$, which is significant for embedded systems and
other resource-tight applications. The time complexity is same as that of SUF,
i.\,e. $O(HW\alpha(HW))$.
}\textru{
Отметим, что SUF2 решает задачу подсчёта СК, используя всего $O(W)$
дополнительной памяти.
}

@@

\texten{
This modification of SUF was implemented to compare the optimization we present
with another optimization suggested by Wu, Otoo and Suzuki\cite{wuotoo}, namely
decision tree. Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not
applicable for condensed image, as it exploits similar observations on
8-connectivity. So this approach is the direct competitor to our optimizations
of SUF.
}\textru{
Эта модификация SUF была добавлена чтобы сравнить нашу оптимизацию с
одной из оптимизаций, предложенных Wu, Otoo и Suzuki\cite{wuotoo}, а именно с
деревом принятия решения.
К сожалению, DTSUF (Decision Tree Scan with Union-Find) не применим для сжатого
изображения, так как использует свойства 8-связности, которыми наша новая
связность не обладает.
}

@@

\texten{
For memory-tight applications such as embedded systems it can be useful to
utilise the suggested condensation in a lazy way, so that without actually
constructing the condensed image we use a function (``view'') calculating color
of each cell with 4 lookups to the original binary image. This is probably most
interesting when combined with SUF2 algorithm, as it is a way to win some time
while avoiding memory footprint increase. However this lazy approach increases
the number of image lookups.
}\textru{
В некоторых случаях может быть более приемлемо использовать сжатие неявно, то
есть вместо построения 
}

@@

So benchmarks were run for all $d \in \set{2\%, 5\%, 10\%, 15\%, \ldots, 90\%,
95\%, 98\%}$ to achieve a picture of how the speedup depends on the number of
object pixels in the image.

It is worth noting again that for ``2x2 copy'' optimization, the apparent time
overhead of constructing the extra condensed image was included into the total
run time, since we measured the total run time of the ``black box'' CC counting
subroutine, supplied with only the original image, so that all the needed
preparations are done inside the subroutine.

@@

The overall benchmark plot of average run time for all algorithms and
optimizations is presented on \Cref{fig:overall}. The actual measured points are
known for sure only for the mentioned values of d, the points are connected to
help distinguish groups of points corresponding to same algorithms.

@@

The worst-case and best-case speedups are shown on the \autoref{tab:speedup}.
Here we see that while ``2x2 view'' optimization is more memory-efficient than
``2x2 copy'', it can slow down the original method in some situations, and
generally it does not give as much performance advantage.

@@

\begin{table}
  \centering
  \begin{tabular}{l|c|c|c|c}
              &               & worst speedup &              & best speedup \\
    Algorithm & worst speedup & density       & best speedup & density \\
    \hline
    \input{tables/speedup}
  \end{tabular}
  \caption{The worst and the best speedups relative to original methods. The
    shown speedup of DTSUF is relative to SUF.}
  \label{tab:speedup}
\end{table}

@@

We present an efficient way to convert the initial binary image with
8-connectivity into a condensed (non-binary) image with new connectivity that is
2 times smaller in each dimension than the initial image. Most approaches for
connected components labeling are still valid for the new image. As time and
memory consumptions of conventional approaches usually depend linearly on the
number of pixels, running them on the condensed image can be more efficient than
running on the initial image. The provided benchmark on random images showed
worst-case speedup factor of $1.7$ for one of our optimizations of the common
SUF approach to CC labeling. The same algorithm is shown to outperform the
previously known decision tree optimization of SUF by at least $1.5$ factor for
image densities less than 80\%.
